Typed Genetic Programming over Lambda Calculus

### 2014-04-29 tex/wds14/wds14.tex
wds14 zaklad

\documentclass{sig-alternate}

\usepackage[utf8]{inputenc}
\usepackage{xspace}
\usepackage[ampersand]{easylist}
\usepackage{qtree}
\usepackage{color}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}

\newenvironment{lizt}
{\begin{easylist}[itemize]}
{\end{easylist}}

\newcommand{\Lets}{Let us\xspace}
\newcommand{\lets}{let us\xspace}
\newcommand{\lterm}{$\lambda$-term\xspace}
\newcommand{\lterms}{$\lambda$-terms\xspace}
\newcommand{\lhead}{$\lambda$-head\xspace}
\newcommand{\lheads}{$\lambda$-heads\xspace}
\newcommand{\la}{\leftarrow\xspace}
\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
\newcommand{\atTree}{@-tree\xspace}
\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
\newcommand{\lh}[1]{\lambda #1}
\newcommand{\sexprTree}{sexpr-tree\xspace}
\newcommand{\SexprTree}{Sexpr-tree\xspace}
\newcommand{\then}{\Rightarrow\xspace}
\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
\newcommand{\ST}{\mathop{\mathrm{ST}}}
\newcommand{\FV}{\mathop{\mathrm{FV}}}
\newcommand{\Scomb }{\mathbf{S}}
\newcommand{\Kcomb }{\mathbf{K}}
\newcommand{\Icomb }{\mathbf{I}}
\newcommand{\bbarr}{\twoheadrightarrow_\beta}
\newcommand{\barr}{\rightarrow_\beta}
\newcommand{\beq}{=_\beta}
\newcommand{\eearr}{\twoheadrightarrow_\eta}
\newcommand{\earr}{\rightarrow_\eta}
\newcommand{\eeq}{=_\eta}
\newcommand{\bearr}{\rightarrow_{\beta\eta}}
\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
\newcommand{\beeq}{=_{\beta\eta}}
\newcommand{\etar}{\twoheadrightarrow_\eta}
\newcommand{\ered}{$\eta$-reduction\xspace}
\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
\newcommand{\bredex}{$\beta$-redex\xspace} 
\newcommand{\lnf}{\textit{lnf}\xspace}
\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
\newcommand{\Bcomb }{\mathbf{B}}   
\newcommand{\BBcomb }{\mathbf{B*}}
\newcommand{\Ccomb }{\mathbf{C}}   
\newcommand{\CCcomb }{\mathbf{C'}}
\newcommand{\SScomb }{\mathbf{S'}}
\newcommand{\ar}{\rightarrow\xspace}
\newcommand{\T}{\mathbb{T}\xspace}
\newcommand{\C}{\mathbb{C}\xspace}
\newcommand{\Real}{\mathbb{R}}
\newcommand{\gar}{\longmapsto}

@@

\newcommand{\Lpr}{\Lambda^\prime}
\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
\newcommand{\ro}[1]{{\color{blue} #1}}
\newcommand{\tom}[1]{{\color{ForestGreen} #1}}
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\nehodi}[1]{{\color{yellow} #1}}

@@

\conferenceinfo{GECCO'14,} {July 12-16, 2014, Vancouver, BC, Canada.}
\CopyrightYear{2014}
\crdata{TBA}
\clubpenalty=10000
\widowpenalty = 10000

\title{Utilization of Reductions and Abstraction Elimination in Typed Genetic Programming}

\numberofauthors{2}

\author{
% 1st. author
\alignauthor
Tom\'{a}\v{s} K\v{r}en\\
  \affaddr{Faculty of Mathematics and Physics}\\
  \affaddr{Charles University in Prague}\\
  \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,}\\
  \affaddr{Prague, Czech Republic}\\
  \email{tomkren@gmail.com}
% 2nd. author
\alignauthor
Roman Neruda\\
  \affaddr{Institute of Computer Science}\\
  \affaddr{Academy of Sciences of the Czech Republic}\\
  \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,}\\
  \affaddr{Prague, Czech Republic}\\
  \email{roman@cs.cas.cz}
}

%\date{24 January 2014}

\maketitle
\begin{abstract}
Lambda calculus representation of programs offers a more expressive alternative to traditional S-expressions. In this paper we discuss advantages of this representation coming from use of reductions (beta and eta) and how to overcome disadvantages caused by variables occurring in the programs by use of the abstraction elimination algorithm. We discuss the role of those reductions in the process of generating initial population and compare several crossover approaches including novel approach to crossover operator based both on reductions and abstraction elimination. The design goal of this operator is to turn the disadvantage of abstraction elimination  - possibly quadratic increase of program size - into a virtue; our approach leads to more crossover points. At the same time, utilization of reductions provides offspring of small sizes.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{TODO}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{TODO}

\keywords{TODO, ToDo, todo}

@@

Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}. Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.   
Use of types naturally opens door to enriching S-expressions,
the traditional GP representation of individuals, with concepts from
lambda calculus, which is simple yet powerful functional, mathematical and programming language extensively used in type theory. Such attempts has shown to be successful \cite{yu01}.

\red{ 
Reprezentace v klasickym kozovy je S-expression.
Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.

Zdá se že použití kombinátorů má řadu výhod (vymenovat nnějaký co jmenuje briggs), 
ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
taky za zkoužku, což podporují nasledující argumenty:

  (a) motivace pro práci s lambda termama je, že za nima košatá teorie, 
      která například popisuje redukce lambda termů (ty zajišťujou jak 
      zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, 
      díky tomu, že se různé stromy redukují na ten samej).
      (ALT znění: 
      pohovořit o tom, že normální formy zmenšujou prohledávaný prostor, 
      protže stačí brát v potaz třídy ekvivalence a ne zamotný termy 
      - reprezentanti těchto tříd ekvivalence jsou pak právě normální formy. )
      (ALT2: 
      Dost podstatný ospravedlnění lambda reprezentace je fakt, 
      že redukce (beta, eta) šlapou právě jen při tom, když máme lambdy.)

  (b) úspornost - abstrakce eliminací produkuje 
      termy až kvadraticky větší než je původní 
      term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
      míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
      podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí do
      takto vnořeného podtermu složitě dopravit).

  (c) použití kombů vyžaduje silnější typovej sytém - 
      kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
      lambdy v pohodě běžej na simply typed lc.

  (d) použití proměnných je typické pro lidi, proto by se mohlo sát že
      i programy generovaný systémem kterej s proměnejma počitá budou o něco 
      "lidštější"

Since abstr-elim a redukce neinvolvujou moc typově rilejtlý problémy, je ok celou věc demostrovat na jednoduchém typovém systému jakám je simply typed lc. Pro more soffisticated type systems would be situation similar.

Hlavním problémem lambda reprezentac proti kombinátorový je 
jak se vypořádat s problémy zpusobenými použitím 
localních proměných (aka jak křížit).

}

%\nehodi{The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation. During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies. Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus. Thus, the natural idea is to employ an exhaustive systematic search. On the other hand, if we were to mimic the standard GP approach, a quite arbitrary yet common and successful ramped half-and-half generating heuristic~\cite{fg} should probably be used. These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies. This design goal also differentiate our approach from the three state of the art proposals for typed GP known to us that are discussed in the following section. Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search. Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.}

\red{ve shrnutí paperu říct, že jeho tělo je tvořeno dvěma částma,
první se zabejvá generovánim a jeho vztahem k redukcím, druhá řeší křížení
s velkym důrazem na vyřešení problémů s lokálníma proměnejma.}

The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions. Main original results about search strategies in individual generating are described in section~\ref{approach}. Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.

\subsection{\red{Randombits}}
\red{

- někam poznamenat že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP..

}

@@

\red{TODO zatim jen zkopčený z minula, opravit podle novýho kontextu,
možná přesunout velkou obhajobu z úvodu sem..}

Yu presents a GP system utilizing
polymorphic higher-order functions\footnote{Higher-order 
function is a function taking another function as 
input parameter.} and lambda abstractions  \cite{yu01}.
Important point of interest in this work is use of
\texttt{foldr} function as a tool for \textit{implicit recursion},
i.e. recursion without explicit recursive calls. 
The terminal set for constructing lambda abstraction subtrees 
is limited to use only constants and variables of that particular
lambda abstraction, i.e., outer variables are not allowed to be used
as terminals in this work. This is significant difference from our approach 
since we permit all well-typed normalized \lterms. From this difference also
comes different crossover operation. We focus more on term generating process; 
their term generation is performed in a similar way as the standard one, 
whereas our term generation also tries to utilize techniques of systematic enumeration.

Briggs and O’Neill present technique 
utilizing typed GP with combinators \cite{kes}.
The difference between approach presented in this work
and our approach is that in this work terms are generated
straight from \textit{library} of combinators and no lambda abstractions
are used. They are using more general polymorphic type system than us
-- the Hindley–Milner type system. They also discuss the 
properties of exhaustive enumeration of terms and compare it with GP search.  
They also present interesting concept of \textit{Generalized
genetic operator} based on term generation.

Binard and Felty use even 
stronger type system (\textit{System F}) \cite{binard2008genetic}.  
But with increasing power of the type system comes increasing difficulty of term generation.
For this reason evolution in this work takes interesting and nonstandard shape 
(fitness is associated with \textit{genes} which are evolved together with \textit{species}
which together participate in creation of individuals).
This differs from our approach, which tries to be generalization of
the standard GP\cite{koza92}.

@@

\section{Preliminaries}
\label{preliminaries}

In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
First, \lets describe a programming language, 
in which the GP algorithm generates individual programs --- the so called \lterms.

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	 	
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}

\textit{Function application} and 
\textit{$\lambda$-abstraction} are concepts
well known from common programming languages. 
For example in JavaScript 
$(M~N)$ translates to expression \texttt{$M$($N$)} and
$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
In other words, the function application 
corresponds to the act of supplying a function 
with an argument, and
the $\lambda$-abstraction is equivalent to 
\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
they were recently introduced to C++, and they are expected to be supported in Java 8.}.

For better readability, 
$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for
$(\dots((M_1~M_2)~M_3)~\dots~M_n)$
and $\lam{x_1 x_2 \dots x_n }{M}$ for 
$\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.

\subsection{$\beta$-reduction}

In order to perform computation there must be some
mechanism for term evaluation. In $\lambda$-calculus there
is $\beta$-reduction for this reason.\\

\newcommand{\bRedex}{$\beta$-redex\xspace}
\newcommand{\bRedexes}{$\beta$-redexes\xspace}
\newcommand{\bArrow}{\rightarrow_\beta\xspace}
\newcommand{\eArrow}{\rightarrow_\eta\xspace}
\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}

A term of a form $\lamb{x}{M}N$ is called \textit{\bRedex}.
A \bRedex can be $\beta$-reduced to term $M[x:=N]$. 
This fact is written as \textit{relation} $\bArrow$ 
of those two terms:
\begin{equation} \label{eq:bRed}
\lamb{x}{M}N \bArrow M[x:=N]
\end{equation}
It is also possible to reduce \textit{subterm \bRedexes} 
which can be formally stated as:
\begin{align*}
P \bArrow Q &\then (R~P)      \bArrow (R~Q) \\
P \bArrow Q &\then (P~R)      \bArrow (Q~R) \\
P \bArrow Q &\then \lam{x}{P} \bArrow \lam{x}{Q}  
\end{align*}

In other words, $\beta$-reduction is the process 
of insertion of arguments supplied to a function into 
its body. \\

Another useful relations are $\bbarr$ and $\beq$ defined as follows.

\begin{enumerate}
 \item \begin{enumerate}
 	\item $M \bbarr M$
 	\item $M \barr N \then M \bbarr N$
 	\item $M \bbarr N , N \bbarr L \then M \bbarr L$ 	
 \end{enumerate}
 \item \begin{enumerate}
 	\item $M \bbarr N \then M \beq N$
 	\item $M \beq N \then N \beq M$
 	\item $M \beq N , N \beq L \then M \beq L$
 \end{enumerate}

\end{enumerate}

We read those relations as follows.
\begin{enumerate}
 	\item $M \bbarr N$ --- "$M$ $\beta$-reduces to $N$."  
 	\item $M \barr N$  --- "$M$ $\beta$-reduces to $N$
 	      in one step."
 	\item $M \beq N$ --- "$M$ is $\beta$-convertible to $N$."	
 \end{enumerate}

\subsection{$\eta$-reduction}

Similarly as for $\beta$-reduction we can define $\eta$-reduction 
except that instead of \ref{eq:bRed} we use:

$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$~

Analogically, a term of a form $\lamb{x}{(M~x)}$ is called 
\textit{$\eta$-redex}.

Relation $\bearr\;=\;\barr \cup \earr$. 
(Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.)

Similarly as for $\bbarr$ and $\beq$ we can define relations 
$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.

\subsection{$\eta^{-1}$-reduction}

$\eta^{-1}$-reduction (also called $\eta$-expansion) is 
the reduction converse to $\eta$-reduction.
Again it may be obtained by replacing \ref{eq:bRed}, now with:

$$M \eeArrow \lamb{x}{(M~x)} \textbf{ ~~~~if } x \not\in FV(M) $$

\subsection{Normal forms}

~\begin{easylist}[enumerate]
& A \lterm is a \textit{$\beta$-normal form} (\bnf) 
if it does not have a $\beta$-redex as subterm.
& A \lterm M \textit{has} a \bnf if $M \beq N$
and $N$ is a \bnf.\\
\end{easylist}
A normal form may be thought of as a result of a term evaluation.

Similarly we can define \enf and \benf.

\subsection{Types etc}

A \lterm as described above
corresponds to a program expression with no type information
included. Now we will describe \textit{types} (or \textit{type terms}).

\begin{definition}
Let $A$ be set of {\it atomic type names}. 
Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
\begin{align*}
\alpha      \in A  &\then   \alpha \in \T \\
\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
\end{align*}
\end{definition}

Type $\sigma \ar \tau$ is type for functions taking as input
something of a type $\sigma$ and returning 
as output something of a type $\tau$. 
$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
combining the previously defined \textit{\lterms} and \textit{types} together.

\begin{definition}~

\begin{enumerate}
 \item 	Let $\Lambda$ be set of {\it \lterms}. 
	Let $\mathbb{T}$ be set of {\it types}.       
	A {\it statement} $M : \sigma$ is a pair 
	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
	Statement $M : \sigma$ is vocalized as 
	{\it "$M$ has type $\sigma$"}.
	The term $M$ is called the {\it subject} of the 
	statement $M : \sigma$.
 \item A \textit{declaration} is a statement 
 $x : \sigma$ where $x \in V \cup C$.

 \item A \textit{context} 
 is set of declarations with distinct variables as subjects.
\end{enumerate}
\end{definition}
%~\\

Context is a basic type theoretic concept suitable as a typed alternative
for terminal and function set in standard GP. 
Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
such that $\Gamma$ does not contain any declaration with $x$ as subject.
We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.

\begin{definition}
A statement $M\colon\sigma$ is \textit{derivable from}
a context $\Gamma$ (notation 
\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
if it can be produced by the following rules.
\begin{align*}
x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
\tur{\Gamma,x:\sigma}{M}{\tau}
&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
\end{align*}
\end{definition}

Přeformulovat aby se to hodilo do tohodle kontextu....
Our goal in term generation is to produce terms $M$
for a given pair $\ul{\tau}{\Gamma}$
such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.

\subsection{Long normal form}
\label{lnf}

%v barendrechtovi je to v definici lnf
%$\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$ místo $M_m$ 
%ale to musí bejt určitě překlep...

\begin{definition}
Let \GMS where 
$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha, n \geq 0$.
	\begin{enumerate}
	  \item	
		Then $M$ is in \textit{long normal form} (\lnf) if following 
		conditions are satisfied.
		\begin{enumerate}
		 \item $M$ is term of the form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_m}$\\
		  (specially for $n = 0$, $M$ is term of the form $f$).
		 \item Each $M_i$ is in \lnf.
		\end{enumerate}	
	  \item 
	    $M$ has a \lnf if $M =_{\beta\eta} N$ and $N$ is in \lnf.
	\end{enumerate}
\end{definition}~

As is shown in \cite{barendregt10}, \lnf has following nice properties.

\begin{proposition}
If $M$ has a \bnf, 
%which according to Theorem 2B.4 is always the case, 
then it also has a unique \lnf, 
which is also its unique \beenf.
\end{proposition}

\begin{proposition}
Every $B$ in \bnf has a \lnf 
$L$ such that $L \twoheadrightarrow_{\eta} B$.
\end{proposition}

\subsection{Abstraction elimination}

\textit{Abstraction elimination} is a process of transforming 
an arbitrary \lterm into \lterm that contains no lambda abstractions
and no bound variables.
The newly produced \lterm may contain function applications, 
free symbols from former \lterm and some new symbols standing for 
combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\

Those combinators are defined as:
\begin{align*}
\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
\Kcomb &= \lam{x\,y}{x} \\
\Icomb &= \lam{x}{x} 
\end{align*}

\Lets describe transformation $\Ae$ performing this 
process.\footnote{In our implementation we must also deal with types, because 
our individuals are annotated with types. But since this process
is fairly straightforward but cumbersome to describe, we will 
skip explaining it here. Reader interested in seeing it can see 
the source code of our implementation. }
\begin{align*}
\Ae[x]           &= x &\\[0.4em]
\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
\Ae[\lam{x}{x}]  &= \Icomb &\\
\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
&\textbf{if } x \in \FV(M)\\
\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
&\textbf{if } x \in \FV(M)\\
&&\vee~ x \in \FV(N)\\
\end{align*}

This is simple version of this process. More optimized version,
in the means of the size of resulting term and its performance
is following one, presented in \cite{jones87}.\\

As is stated in \cite{jones87},
the biggest disadvantage of this technique is that the translated
term is often much larger than in its lambda form --- the size of
the translated term can be proportional to the
square of the size of the original term.

But the advantage is also tempting --- no need to deal with variables
and lambda heads.

\section{Term generating and reductions}

\red{
V této sekci prodiskutujeme generování termův lnf, kteroužto můžeme chápat jako
příme zobecnění S-expressions do lambda termů. Povíme si jaký vztah takové termy mají
k beta a eta redukci, jaká je jejich výhoda a jak je lze jednoduše převést
do úspornější beta eta nf. 
}

\subsection{Grammar producing \lterms in \lnf}

In \cite{barendregt10} is shown term generating grammar with 
following rules.

>>>>

%to be compiled with LATEX 2e

%use this setting with 11pt basic font:
\documentclass[11pt]{article}
\usepackage{wds11,epsf}

%use this setting with 10pt basic font:
%\documentclass{article}
%\usepackage{wds10,epsf}

%use to get author-year citations with BibTeX
\usepackage[square]{natbib}

\usepackage{xspace}

\newcommand{\Lets}{Let us\xspace}
\newcommand{\lets}{let us\xspace}
\newcommand{\lterm}{$\lambda$-term\xspace}
\newcommand{\lterms}{$\lambda$-terms\xspace}
\newcommand{\lhead}{$\lambda$-head\xspace}
\newcommand{\lheads}{$\lambda$-heads\xspace}
\newcommand{\myla}{\leftarrow\xspace}
\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
\newcommand{\atTree}{@-tree\xspace}
\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
\newcommand{\lh}[1]{\lambda #1}
\newcommand{\sexprTree}{sexpr-tree\xspace}
\newcommand{\SexprTree}{Sexpr-tree\xspace}
\newcommand{\then}{\Rightarrow\xspace}
\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
\newcommand{\ST}{\mathop{\mathrm{ST}}}
\newcommand{\FV}{\mathop{\mathrm{FV}}}
\newcommand{\Scomb }{\mathbf{S}}
\newcommand{\Kcomb }{\mathbf{K}}
\newcommand{\Icomb }{\mathbf{I}}
\newcommand{\bbarr}{\twoheadrightarrow_\beta}
\newcommand{\barr}{\rightarrow_\beta}
\newcommand{\beq}{=_\beta}
\newcommand{\eearr}{\twoheadrightarrow_\eta}
\newcommand{\earr}{\rightarrow_\eta}
\newcommand{\eeq}{=_\eta}
\newcommand{\bearr}{\rightarrow_{\beta\eta}}
\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
\newcommand{\beeq}{=_{\beta\eta}}
\newcommand{\etar}{\twoheadrightarrow_\eta}
\newcommand{\ered}{$\eta$-reduction\xspace}
\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
\newcommand{\bredex}{$\beta$-redex\xspace} 
\newcommand{\lnf}{\textit{lnf}\xspace}
\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
\newcommand{\Bcomb }{\mathbf{B}}   
\newcommand{\BBcomb }{\mathbf{B*}}
\newcommand{\Ccomb }{\mathbf{C}}   
\newcommand{\CCcomb }{\mathbf{C'}}
\newcommand{\SScomb }{\mathbf{S'}}
\newcommand{\ar}{\rightarrow\xspace}
\newcommand{\T}{\mathbb{T}\xspace}
\newcommand{\C}{\mathbb{C}\xspace}
\newcommand{\Real}{\mathbb{R}}

@@

\newcommand{\Lpr}{\Lambda^\prime}
\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
\newcommand{\ro}[1]{{\color{blue} #1}}
\newcommand{\tom}[1]{{\color{ForestGreen} #1}}
\newcommand{\red}[1]{{\color{red} #1}}

%\tighten

% Preamble Information

%%%%%%%%%%% nevim co to tu bylo za omg, tak jsem to zakomentoval
%%\lefthead{SAFRANKOVA ET AL.}
%%\righthead{MAGNETOSHEATH RESPONSE}

\setcounter{secnumdepth}{0}

@@

\title{Typed Genetic Programming over Lambda Calculus}

\author{T. K\v ren}

\affil{Charles University, Faculty  of  Mathematics  and  Physics,
     Prague, Czech Republic.}

\begin{abstract}
Todo.
\end{abstract}

\begin{article}

@@

@@

\cite{yu01} presents a GP system utilizing
polymorphic higher-order functions\footnote{Higher-order 
function takes another function as an input parameter.
} and lambda abstractions.
Important point of interest in this work is use of \texttt{foldr}\footnote{ In the functional programming language Haskell \texttt{foldr} can be defined as:\\ \texttt{foldr f z [] $~~~$  = z\\ 
foldr f z (x:xs) = f x (foldr f z xs) }} function as a tool for \textit{implicit recursion},
i.e. recursion without explicit recursive calls. 
The terminal set for constructing lambda abstraction subtrees 
is limited to use only constants and variables of that particular
lambda abstraction, i.e., outer variables are not allowed to be used
as terminals in this work. This is significant difference from our approach 
since we permit all well-typed normalized \lterms. From this difference also
comes different crossover operation. We focus more on term generating process; 
their term generation is performed in a similar way as the standard one, 
whereas our term generation also tries to utilize techniques of systematic enumeration.

\cite{kes} present technique 
utilizing typed GP with combinators.
The difference between approach presented in this work
and our approach is that in this work terms are generated
straight from \textit{library} of combinators and no lambda abstractions
are used. They are using more general polymorphic type system than us
-- the Hindley–Milner type system. They also discuss the 
properties of exhaustive enumeration of terms and compare it with GP search.  
They also present interesting concept of \textit{Generalized
genetic operator} based on term generation.

\cite{binard2008genetic} use even 
stronger type system (\textit{System F}).  
But with increasing power of the type system comes increasing difficulty of term generation.
For this reason evolution in this work takes interesting and nonstandard shape 
(fitness is associated with \textit{genes} which are evolved together with \textit{species}
which together participate in creation of individuals).
This differs from our approach, which tries to be generalization of
the standard GP\cite{koza92}.

@@

%%% Abych věděl jak citovat tak sem si tu kousek nechal
%The interaction of the solar wind with the Earth's magnetosphere
%generates a population of backstreaming ions directed from the
%bow shock into the solar wind. This high-energy population was
%invoked to explain Hot Flow Anomalies (HFAs)
%\citep[e.g.][]{sw}. \citet{th} identified as heated regions of solar wind plasma
%flowing nearly perpendicular to the Earth-Sun line.
%The main observational features of HFAs include \citep[and others]{sw}: (1) Central regions with hot plasma flowing
%significantly slower than that in the ambient solar wind in a
%direction highly deflected (nearly $90^o$) from the Sun-Earth line.
%The flow velocities are often roughly tangential to the nominal
%bow shock shape \citep{sw}.
%(2) HFAs are bounded by regions of enhanced
%magnetic field strength, density, and temperature. The outer
%edges of these enhancements are fast shocks generated by
%pressure enhancements within the core region. The inner
%edges of the enhancements are probably  tangential
%discontinuities.
%Published examples indicated that  many HFAs are bounded by only
%one enhancement.
%(3) HFAs occur in conjunction with significant changes in the IMF
%direction. The angle between pre- and post event orientations is
%typically $\sim 70^o$.

Todo.

\section{Conclusion}
Todo.

\acknowledgments %\footnotesize
{???}

%% TWO METHODS FOR INCLUDING THE BIBLIOGRAPHY (LIST OF REFERENCES)
%% EITHER TYPE IN THE ENTRIES YOURSELF AS SHOWN HERE IN
%% `thebibliography' ENVIRONMENT,
%%        OR
%% USE THE FOLLOWING TWO COMMANDS SO THAT BIBTEX WILL GENERATE
%% `thebibliography' TEXT FOR YOU AND READ IT IN.
%%
\bibliographystyle{egs}%<-- LIST OF REFERENCES TO BE IN "EGS" STYLE (FILE "EGS.STY")
\bibliography{my}       %<-- REFERENCES ARE IN FILE "SAMPLE.BIB"
%%
%%  IF THE ABOVE TWO COMMANDS ARE USED, THEN thebibliography ENVIRONMENT
%%  MUST BE REMOVED.

%\begin{thebibliography}{}
%
%\itemsep=0.4ex
%%\footnotesize
%\bibitem[\it Klimov and Friend(1997)]{kl}
%Klimov, S., Friend A., ASPI experiment: Measurements of fields and
%waves onboard the Interball-1 spacecraft, {\it Ann. Geophys.}, 15,
%514-527, 1997.
%
%\bibitem[\it Kudela et al.(1995)]{ku}
%Kudela, K., M. Slivka, J. Rojko, and V. N. Lutsenko, The
%apparatus DOK-2 (project INTERBALL): Output data structure and
%modes of operation, preprint of Inst. Exp. Phys. UEF-01-95,
%Kosice, 20, 1995.
%
%\bibitem[\it Lin(1997)]{li}
%Lin Y., Generation of anomalous flows near the bow shock by
%its interaction with interplanetary discontinuities, {\it J. Geophys.
%Res.}, 102, 24265-24281, 1997.
%
%\bibitem[\it Lutsenko et al.(1995)]{lu}
%Lutsenko, V. N., J. Rojko, K. Kudela, T. V. Gretchko, J. Balaz,
%J. Matisin, E. T. Sarris, K. Kalaitzides, and N. Paschalidis,
%Energetic particle experiment DOK-2 (Interball project), in:
%{it INTERBALL Mission and Payload,} ed. by Yu. Galperin, IKI-CNES,
%249, 1995.
%
%
%\bibitem[\it Onsager et al.(1990a)]{ona}
%Onsager, T. G., M. F. Thomsen, J. G. Gosling, and S. J.
%Bame, Observational test of a hot flow anomaly formation
%mechanism, {\it J. Geophys. Res.}, 95, 11.967-11.974, 1990a.
%
%\bibitem[\it Onsager et al.(1990b)]{onb}
%Onsager, T. G., M. F. Thomsen, and D. Winske,
%Hot flow anomaly formation by magnetic deflection,
%{\it Geophys. Res. Lett.}, 17, 1621-1624, 1990b.
%
%\bibitem[\it Schwartz et al.(1999d)]{sw}
%Schwartz, S. J., G. E. Paschmann, N. Sckopke, T. M. Bauer, M. W.
%Dunlop, A. N. Fazakerley, and M. F. Thomsen,  Hot flow anomalies
%revisited, {\it Int. J. Geomag. and Aeronomy}, 1, 1999, in print.
%
%\bibitem[\it Thomsen et al.(1992)]{th}
%Thomsen M. F., J. G. Gosling, and S. J.
%Bame, Observational test of a hot flow anomaly formation
%mechanism, {\it J. Geophys. Res.}, 97, 1967-1974, 1992.
%
%
%\end{thebibliography}

\end{article}
\end{document}
